High-resolution ALMA observations of V4046\,Sgr: a circumbinary disc with a thin ring

### 2020-11-20 Draft_V4046Sgr_mnras.tex
Updates from Overleaf

Assuming a three dimensional model in a cylindrical reference frame with coordinates ($r$, $\theta$, $z$), the gas density ($\rho_{\mathrm{gas}}$) distribution follows
\begin{equation}
  \rho_{\mathrm{gas}}(r,z) =\frac{\Sigma(r) \,\delta(r)}{\sqrt{2\pi} \, H(r)} \mathrm{exp}\left[-\frac{1}{2} \left(\frac{z}{H(r)}\right)^2\right],
\end{equation}
where $H(r)$ is the scale height profile and $\Sigma(r)$ is the surface density profile. $\Sigma(r)$ is defined by a combination of four individual components: the inner disc, Ring6, Ring13 and Ring24.
\begin{equation}
  \Sigma(r) = \Sigma_{\mathrm{inner\,disc}} + \Sigma_{\mathrm{Ring6}} + \Sigma_{\mathrm{Ring13}} + \Sigma_{\mathrm{Ring24}}
\end{equation}

%$\delta(r)$ parameterises the density drops in the gaps and cavities,

The inner disk follows a power law function defined by
\begin{equation}
  \Sigma_{\mathrm{inner\,disc}}(r) =\Sigma_\mathrm{c} \left(\frac{r}{R_\mathrm{c}}\right)^{-\gamma}  \, \mathrm{exp}\left[-\left(\frac{r}{R_\mathrm{c}}\right)^{2-\gamma}\right],
\end{equation}
where $R_c$ is a characteristic radius and $\gamma$ is the surface density gradient. We used $R_c$ = 16 and a fixed $\gamma$ = 1 as it is a typical value for discs \citep{Andrews_2009,Andrews_2010}.

We choose to use a Gaussian profiles to parameterize Ring6 and Ring13,
\begin{equation}
  \Sigma_{\mathrm{Ring}}(r) = \frac{\Sigma_\circ}{\sqrt{2 \pi} \sigma}
  \, \mathrm{exp}\left[-\frac{1}{2}\left(\frac{r-\mu}{\sigma}\right)^{2}\right],
\end{equation}

%The $\delta(r)$ function, that account for deviations from the power law, is defined as:
%\begin{align}
%\delta(r) = \left\{ \begin{array}{cc} 
%                 5E-6 & \hspace{5mm} 0.3<r<5 \\
%                 5E-3 & \hspace{5mm} 5<r<8  \\
%                 5E-5 & \hspace{5mm} 8<r<12\\
%                 2.5E-2 & \hspace{5mm} 12<r<19.5 \\
%                 1E-8 & \hspace{5mm} 19.5<r<20.5 \\
%                 1E-4 & \hspace{5mm} 20.5<r<25 \\
%                 1.7E-1 & \hspace{5mm} 25<r<120 \\
%                \end{array} \right.
%\end{align}

The parametric scale height profiles for the gas and for each dust population are 
\begin{equation}
    \label{scale}
  H(r)=\chi \, H_{o} \, [r/r_{o}]^{\psi(r)},
\end{equation}
where $H_\circ$ is the scale height at r = $r_\circ$, $\psi$ is the flaring index and $\chi$ is a scaling factor (in the range $0-1$) that mimics dust settling. For the gas and the small dust grains $\chi$ = 1 and for the millimetre-size dust population we assign a fixed $\chi$ = 1/2 for simplicity, as in \citet{Rosenfeld_2013}.

For the vertical structure, \citet{dOrazi} found flaring angles of $\varphi$ = 6.2$\pm$0.6\,$\degr$ for the inner ring and $\varphi$ = 8.5$\pm$1.0\,$\degr$ for the outer one. Our model uses the same values, with two different flaring indexes $\psi$. The separation between the two values were set at $r$ = 19\,au with $\psi=0.2$ for Ring6 and Ring13, and $\psi=0.5$ for Ring24. The scale height is $H_\circ = 0.8775$\,au at $r_\circ = 19.5$\,au.

@@

Since the large dust grains are less coupled to the gas, their distribution has some important differences that require a special parameterisation, with a large inner cavity and scaling factors that account for the larger gap between the rings. For the inner ring, the same profile as the gas is used, but with different values for $R_c$ and $\Sigma_c$. The greatest difference is in the outer ring, where we used a decaying power law modulated by the sum of two Gaussian profiles. In an effort to recreate the break seen in the outer ring between 29.8 and 37.9\,au, we chose a different value for $\gamma$ = -3. For the large dust grains the surface density profile is thus

\begin{equation}
  \Sigma_{\mathrm{large-dust}}(r) = \Sigma_{\mathrm{c}2} \left(\frac{r}{R_{\mathrm{c}2}}\right)^{-\gamma} \exp\left[-\left(\frac{r}{R_{\mathrm{c}2}}\right)^{2-\gamma}\right] .
\end{equation}

The inner part of the outer ring follows a different behaviour. Between $R_\mathrm{in}$ and $R_\mathrm{peak}$, marking the beginning and the maximum density peak of the outer ring respectively, the gas density profile is multiplied by an additional factor
\begin{equation}
    \epsilon(r) = \left(\frac{ r - R_\mathrm{in}}{R_\mathrm{peak} - R_\mathrm{in}}\right)^3.
\end{equation}
$R_\mathrm{in}$ and $R_\mathrm{peak}$ would depend on the dust population, with $R_\mathrm{in}$ = 19.5\,au and $R_\mathrm{peak}$ = 25\,au for the small dust grains and $R_\mathrm{in}$ = 24.2\,au and $R_\mathrm{peak}$ = 26.6\,au for the large dust grains. This parameter allow us to model a smoother inner wall of the outer ring.

@@

The gas and small dust surface density profiles begin at 0.3\,au, out of the zone expected to be cleared by dynamical interactions with the central binary \citep{Art_Lu}, which is expected to be $\sim$0.2\,au. For the small dust grains, we propose a three ringed structure. The innermost ring of the model is located from 5 to 8\,au, this ring, introduced as Ring6 in Sec.~\ref{sec:Observations}, is not detectable in the IRDIS observations but was required to account for the SED. Then follow the two rings observed by IRDIS: Ring13, from 12 to 19.5\,au, %which is surrounded by very narrow 1\,au-wide gap, 
and then Ring24, that starts 19.5\,au and has its maximum density peak at 25\,au. The small-dust density then decays exponentially with radius.

On the other hand, in order to reproduce the double-ringed morphology of the millimetre continuum emission, we required a model with a wide central cavity ($r = 11.7$\,au), a narrow 3.7\,au-wide inner ring (which is Ring13) followed by a 8.2\,au gap that separates the outer ring (Ring24), which starts at 22\,au and reaches peak densities at 27.7\,au. This last ring has an observed break between 30.2 and 36.7\,au, which is reproduced in the model with a discontinuity in the exponent $\gamma$, along with the sum of the decaying exponentials. We also truncated the model large-grain density at $\sim$55\,au.

>>>>

Assuming a three dimensional model in a cylindrical reference frame with coordinates ($r$, $\theta$, $z$), the gas density ($\rho_{\mathrm{gas}}$) distribution follows
\begin{equation}
  \rho_{\mathrm{gas}}(r,z) =\frac{\Sigma(r) \,\delta(r)}{\sqrt{2\pi} \, H(r)} \mathrm{exp}\left[-\frac{1}{2} \left(\frac{z}{H(r)}\right)^2\right],
\end{equation}
where $H(r)$ is the scale height profile and $\Sigma(r)$ is the gas surface density profile.

The parametric scale height profiles for the gas and for each dust population are 
\begin{equation}
    \label{scale}
  H(r)=\chi \, H_{\circ} \, [r/r_{\circ}]^{\psi(r)},
\end{equation}
where $H_\circ$ is the scale height at r = $r_\circ$, $\psi$ is the flaring index and $\chi$ is a scaling factor (in the range $0-1$) that mimics dust settling. For the gas and the small dust grains $\chi$ = 1 and for the millimetre-size dust population we assign a fixed $\chi$ = 1/2 for simplicity, as in \citet{Rosenfeld_2013}.

For the vertical structure, \citet{dOrazi} found flaring angles of $\varphi$ = 6.2$\pm$0.6\,$\degr$ for the inner ring and $\varphi$ = 8.5$\pm$1.0\,$\degr$ for the outer one. Our model uses the same values, with two different flaring indexes $\psi$. The separation between the two values were set at $r$ = 19\,au with $\psi=0.2$ for Ring6 and Ring13, and $\psi=0.5$ for Ring24. The scale height is $H_\circ = 1$\,au at $r_\circ = 19$\,au.

$\Sigma(r)$ begin at 0.3\,au, out of the zone expected to be cleared by dynamical interactions with the central binary \citep{Art_Lu}, which is expected to be $\sim$0.2\,au. We propose a three ringed structure, so we defined the total gas surface density by a combination of four individual components: the inner disc, Ring6, Ring13 and Ring24.
\begin{equation}
  \Sigma(r) = \Sigma_{\mathrm{inner\,disc}} + \Sigma_{\mathrm{Ring6}} + \Sigma_{\mathrm{Ring13}} + \Sigma_{\mathrm{Ring24}}
\end{equation}

The inner disk follows a power law function defined by
\begin{equation}
  \Sigma_{\mathrm{inner\,disc}}(r) =\Sigma_\mathrm{c} \left(\frac{r}{R_\mathrm{c}}\right)^{-\gamma}  \, \mathrm{exp}\left[-\left(\frac{r}{R_\mathrm{c}}\right)^{2-\gamma}\right],
\end{equation}
where $R_c$ is a characteristic radius and $\gamma$ is the surface density gradient. We used $R_c$ = 16\,au, $\Sigma_\mathrm{c} =2.6\times10^{-4}$ and a fixed $\gamma$ = 1 as it is a typical value for discs \citep{Andrews_2009,Andrews_2010}. We chose to use a Gaussian profiles to parameterize Ring6 and Ring13,
\begin{equation}
  \Sigma_{\mathrm{Ring}}(r) = \frac{\Sigma_\circ}{\sqrt{2 \pi} \sigma}
  \, \mathrm{exp}\left[-\frac{1}{2}\left(\frac{r-\mu}{\sigma}\right)^{2}\right],
\end{equation}
where we define constants that correspond to the centroid radii \{$\mu_{\mathrm{Ring6}}=6$\,au, $\mu_{\mathrm{Ring13}}=15$\,au\}, ring widths \{$\sigma_{\mathrm{Ring6}}=1$\,au, $\sigma_{\mathrm{Ring13}}=2.05$\,au\}, and normalizations \{ $\Sigma_{\circ,\mathrm{Ring6}}=1\times10^{13}$, $\Sigma_{\circ,\mathrm{Ring13}}=4.2\times10^{13}$\} for both structure components separately.

For Ring24 we used the same power law than the inner disc but scaled by a $\delta_{\mathrm{sd}}$ factor:
\begin{equation}
  \Sigma_{\mathrm{Ring24}} = \Sigma_{\mathrm{inner\,disc}}(r)\, \delta_{\mathrm{sd}},
\end{equation}
with $\delta_{\mathrm{sd}}$ = 5.2$\times 10^4$. But the inner wall of Ring24 follows a different behaviour. Between $R_\mathrm{in}$ and $R_\mathrm{peak}$, marking the beginning and the maximum density peak of the outer ring respectively, the gas density profile is multiplied by an additional factor
\begin{equation}
    \epsilon(r) = \left(\frac{ r - R_\mathrm{in}}{R_\mathrm{peak} - R_\mathrm{in}}\right)^3.
\end{equation}
This parameter allow us to model a smoother inner wall of the outer ring, with $R_\mathrm{in,gas}$ = 19.\,au and $R_\mathrm{peak,gas}$ = 26.8\,au. This factor is also applied to the large dust population with $R_\mathrm{in,ld}$ = 24.6\,au and $R_\mathrm{peak,ld} = R_\mathrm{peak,gas}$.

@@

Since the large dust grains are less coupled to the gas, their distribution has some important differences that require a special parameterisation, with a large inner cavity, a larger gap between the rings and a break in the outer ring. The large-dust surface density profile is defined by the sum of its two components as For the large dust grains the surface density profile is thus
\begin{equation}
  \Sigma_{\mathrm{ld}}(r) = 
  \Sigma_{\mathrm{Ring13,ld}} + \Sigma_{\mathrm{Ring24,ld}}.
\end{equation}

For the inner ring, we chose a Gaussian profile parameterized with centroid radius $\mu_{\mathrm{Ring13, ld}}=13.36$\,au, ring width of $\sigma_{\mathrm{Ring13,ld}}=0.54$\,au and $\Sigma_{\circ,\mathrm{Ring13,ld}}=4.2\times10^{13}$. For Ring24 we used the same distribution of the gas, but with the difference that, in an effort to recreate the break seen in the outer ring, between 28.6 and 37\,au we chose a different value for $\gamma$ = -5 and $R_c$ = 38. We also truncated the model large-grain density at $\sim$64\,au. Thus, the surface density of the outer ring of large grains dust is defined by
\begin{equation}
    \Sigma_{\mathrm{Ring24,ld}} = \Sigma_{\mathrm{c}} \left(\frac{r}{R_{\mathrm{c}}}\right)^{-\gamma} \exp\left[-\left(\frac{r}{R_{\mathrm{c}}}\right)^{2-\gamma}\right]\,\delta_{\mathrm{ld}}(r),
\end{equation}
where
\begin{equation}
  \delta_{\mathrm{ld}}(r) =
  \begin{cases}
  0                 & r < 24.6\text{\,au} \\
  4.37 \times10^{4} & 24.6 < r < 28.6\text{\,au} \\
  2.04 \times10^{4} & 28.6 < r < 37\text{\,au} \\
  1.82 \times10^{5} & 37   < r < 64\text{\,au} \\
  0                 & r > 64\text{\,au.}
  \end{cases}
\end{equation}

@@
